---
title: 'Sampling Activity and Sensor Performance'
author: "Hallie Rutten"
date: '2022-06-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load libraries -----
library(plotly)
library(ggthemes)

# basic read -----
source('Read-bats.R')

# sensor data read -----
source('Read-sensorDates.R')
  
# join bats and sensor site data -----
bats <- left_join(bats,sensorDates, by=c('siteID','DATE'))
```

# Reminder:

Before you run this code, make sure...

- you have a main folder that contains two subfolders; 'Data' and 'Code'

- the subfolder called 'Data' holds all the data files,
  sorted in the proper place in the folder tree
  
- the subfolder called 'Code' holds all the code files,
  such as this data read, the exploratory markdowns, and the dashboard
  
- you set your session working directory to the 'Code' folder
  ( Top of the RStudio interface: Session > Set Working Directory > Choose Directory )




# EXPLORATORY:

## Sampling Effort Trends

```{r, message=FALSE, warning=FALSE}
# recording effort by date -----
sensor.days <- sensorDates %>% 
  filter( !is.na(project) | !is.na(monitor) ) %>%
  group_by(DATE) %>% 
  summarize( nSensors = n() )

ggplot( sensor.days ) +
  geom_line( aes(x=DATE, y=nSensors) )

ggplot( sensor.days ) +
  geom_point( aes(x=DATE, y=nSensors), alpha=0.4 )
```

## Sensor Performance Trends

```{r, message=FALSE, warning=FALSE}
# monitor accuracy -----
IDs <- unique(bats$AUTO.ID)
IDbats <- IDs[which( IDs!='no.ID' & IDs!='Noise')]

monitor.acc <- bats %>% 
  group_by(monitor,year) %>% 
  summarize( nBats = sum(as.numeric(AUTO.ID %in% IDbats)),
             unID  = sum(as.numeric(grepl("no.ID",AUTO.ID,ignore.case=TRUE))),
             noise = sum(as.numeric(grepl("Noise",AUTO.ID,ignore.case=TRUE))),
             total = sum(nBats,unID,noise),
             p.unID = unID/total, 
             p.noise = noise/total )

ggplot( monitor.acc ) +
  geom_col( aes(x=monitor, y=p.unID, fill=p.unID) ) +
  facet_wrap(~year) +
  theme(legend.position='none') +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title="Monitor 'no ID' observations",
       x="Monitor number", y="proportion of unidentified bats",
       caption="DataLab 2022")

ggplot( monitor.acc ) +
  geom_col( aes(x=monitor, y=p.noise, fill=-p.noise) ) +
  facet_wrap(~year) +
  theme(legend.position='none') +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title="Monitor 'noise' observations",
       x="Monitor number", y="proportion of noise",
       caption="DataLab 2022")
```










# DASHBOARD PREP:

## Sampling Activity

```{r, message=FALSE, warning=FALSE}
selectedYears <- unique(bats$year)

bats.sub <- bats %>% 
  filter( year %in% selectedYears )

sampling.granularity <- 'month'
sampling.style <- 'line'
```

```{r, message=FALSE, warning=FALSE}
    if(sampling.granularity == 'year'){
      sensorDates <- sensorDates %>% mutate( div=year(DATE) )
    } else if(sampling.granularity == 'month'){
      sensorDates <- sensorDates %>% 
        mutate( div=as.Date(paste(year(DATE),month(DATE),1,sep='-')) )
    }else if(sampling.granularity == 'day'){
      sensorDates <- sensorDates %>% mutate( div=DATE )
    }
```

```{r, message=FALSE, warning=FALSE}
    sampling.days <- sensorDates %>% 
      filter( !is.na(project) | !is.na(monitor) ) %>%
      group_by(div) %>% 
      summarize( nSites = n() )
```

```{r, message=FALSE, warning=FALSE}
    sample.plot <- ggplot( sampling.days, aes(x=div, y=nSites) )
    
    if(sampling.style == 'line'){
      sample.plot <- sample.plot + geom_line()
    }
    if(sampling.style == 'points'){
      sample.plot <- sample.plot + geom_point(alpha=0.4)
    }
    
    sample.plot +
      labs(title="Sampling Activity",
           x=sampling.granularity, y="N Sites",
           caption="Sewanee Bat Study, DataLab 2022")
```


## Sensor accuracy

```{r, message=FALSE, warning=FALSE}
selectedYears <- unique(bats$year)
selectedComps <- unique(bats$COMPARTMENT)

bats.sub <- bats %>% 
  filter( year %in% selectedYears,
          COMPARTMENT %in% selectedComps )

sensor.wrapVar <- 'year'
sensor.ob <- 'Noise'
```

```{r, message=FALSE, warning=FALSE}
if(sensor.wrapVar == 'year'){
  bats.monitor <- bats.sub %>% mutate( wrapV=year)
}

if(sensor.wrapVar == 'none'){
  bats.monitor <- bats.sub %>% 
      group_by(monitor) %>% 
      summarize( nBats = sum(as.numeric(!grepl('no.ID|Noise',AUTO.ID))),
                 unID  = sum(as.numeric(grepl('no.ID',AUTO.ID))),
                 noise = sum(as.numeric(grepl("Noise",AUTO.ID))),
                 total = n(),
                 p.unID = unID/total, 
                 p.noise = noise/total )
} else {
  bats.monitor <- bats.monitor %>% 
      group_by(monitor,wrapV) %>% 
      summarize( nBats = sum(as.numeric(!grepl('no.ID|Noise',AUTO.ID))),
                 unID  = sum(as.numeric(grepl('no.ID',AUTO.ID))),
                 noise = sum(as.numeric(grepl("Noise",AUTO.ID))),
                 total = n(),
                 p.unID = unID/total, 
                 p.noise = noise/total )
}
```

```{r, message=FALSE, warning=FALSE}
if(sensor.ob == 'Noise'){
  monitor.p <- ggplot( bats.monitor, aes(x=monitor, y=p.noise, fill=-p.noise) )
} else {
  monitor.p <- ggplot( bats.monitor, aes(x=monitor, y=p.unID, fill=-p.unID) )
}

monitor.p <- monitor.p +
  geom_col() +
  theme(legend.position='none') +
  theme(axis.text.x = element_text(angle = 90))

if(sensor.wrapVar == 'none'){
  monitor.p
} else {
  monitor.p + facet_wrap(~wrapV)
}
```




## Mic accuracy

```{r, message=FALSE, warning=FALSE}
selectedYears <- unique(bats$year)
selectedComps <- unique(bats$COMPARTMENT)

bats.sub <- bats %>% 
  filter( year %in% selectedYears,
          COMPARTMENT %in% selectedComps )

sensor.wrapVar <- 'year'
sensor.ob <- 'Noise'
```

```{r, message=FALSE, warning=FALSE}
if(sensor.wrapVar == 'year'){
  bats.mic <- bats.sub %>% mutate( wrapV=year)
}

if(sensor.wrapVar == 'none'){
  bats.mic <- bats.sub %>% 
      group_by(mic) %>% 
      summarize( nBats = sum(as.numeric(!grepl('no.ID|Noise',AUTO.ID))),
                 unID  = sum(as.numeric(grepl('no.ID',AUTO.ID))),
                 noise = sum(as.numeric(grepl("Noise",AUTO.ID))),
                 total = n(),
                 p.unID = unID/total, 
                 p.noise = noise/total )
} else {
  bats.mic <- bats.mic %>% 
      group_by(mic,wrapV) %>% 
      summarize( nBats = sum(as.numeric(!grepl('no.ID|Noise',AUTO.ID))),
                 unID  = sum(as.numeric(grepl('no.ID',AUTO.ID))),
                 noise = sum(as.numeric(grepl("Noise",AUTO.ID))),
                 total = n(),
                 p.unID = unID/total, 
                 p.noise = noise/total )
}
```

```{r, message=FALSE, warning=FALSE}
if(sensor.ob == 'Noise'){
  mic.p <- ggplot( bats.mic, aes(x=mic, y=p.noise, fill=-p.noise) )
} else {
  mic.p <- ggplot( bats.mic, aes(x=mic, y=p.unID, fill=-p.unID) )
}

mic.p <- mic.p +
  geom_col() +
  theme(legend.position='none') +
  theme(axis.text.x = element_text(angle = 90))

if(sensor.wrapVar == 'none'){
  mic.p
} else {
  mic.p + facet_wrap(~wrapV)
}
```
